---
title: "Homework 3"
author: "Brandon Leslie"
format: pdf
editor: visual
---

# Forecasting Principles and Practice

```{r}
library(tidyverse)
library(tidymodels)
library(USgas)
library(scales)
library(fpp3)
library(tsibbledata)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Q4: US Gas

### Tsibble

```{r}
tsibble1 <- us_total %>% 
  as_tsibble(
    key = state,
    index = year
  )
glimpse(tsibble1)
```

### Graph

```{r}

NewEngland = c("Maine", "Vermont", "New Hampshire", "Massachusetts", "Connecticut", "Rhode Island")
tsibble1 %>%
  filter(state %in% NewEngland) %>% 
  ggplot(aes(year, y)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess", color = "red") +
  scale_y_continuous(labels = label_number(scale = 0.00001, suffix = "K")) +
  scale_x_continuous(breaks = breaks_width(5)) +
  labs(
    title = "Anual Natural Gas Consumption",
    subtitle = "By state",
    y = "Gas in gallons by Millions",
    x = "Year"
  ) +
  facet_wrap(~state) +
  theme_bw()
  
```

## Q7: Australian Retail Data

```{r}
set.seed(84625314)
myseries <- aus_retail %>%
  as_tsibble() %>% 
  filter(`Series ID` == sample(unique(aus_retail$`Series ID`),1)) %>%
  as_tsibble(index = Month)


autoplot(myseries)
gg_season(myseries)
gg_subseries(myseries)
gg_lag(myseries)
myseries %>% 
  ACF(Turnover) %>% 
  autoplot()
  
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

-   **Seasonality**
    -   There seems to be an increase in turnovers in late fall/early winter, as-well as a peak in the month of march.
-   **Cyclicity**
    -   In the autoplot graph, cyclicity is represented through repeated sudden peaks, then a dip, with.a short after-peak, followed by a slight increase, then this process repeats to another sudden peak.
-   **Trends**
    -   As we progress through time, we can see a positive trend as years & months go by.

## Q8: "Total Private" Timeseries

```{r}
explore_ts <- function(data, index, value) {
  index_sym <- rlang::ensym(index)
  value_sym <- rlang::ensym(value)

  ts_data <- data %>%
    filter(!is.na(!!value_sym)) %>%
    as_tsibble(index = !!index_sym)

  print(autoplot(ts_data, !!value_sym) + ggtitle("Autoplot"))
  print(gg_season(ts_data, !!value_sym) + ggtitle("Seasonal Plot"))
  print(gg_subseries(ts_data, !!value_sym) + ggtitle("Subseries Plot"))
  print(gg_lag(ts_data, !!value_sym) + ggtitle("Lag Plot"))
  print(ts_data |> ACF(!!value_sym) |> autoplot() + ggtitle("ACF Plot"))
}


```

```{r}
# 1. Total Private Employed
private_employed <- us_employment |> filter(Title == "Total Private")
explore_ts(private_employed, Month, Employed)

# 2. Bricks
explore_ts(aus_production, Quarter, Bricks)


# 4. H02 Cost
h02 <- PBS |> filter(ATC2 == "H02") |> summarise(Cost = sum(Cost))
explore_ts(h02, Month, Cost)

# 5. Barrels
explore_ts(us_gasoline, Week, Barrels)

```

### Summary of Time Series Features

| Series      | Seasonality | Cyclicity | Trend    | Unusual Years     |
|-------------|-------------|-----------|----------|-------------------|
| Retail (AU) | Strong      | Light     | Upward   | 2020              |
| Employment  | Mild        | Clear     | Upward   | 2009, 2020        |
| Bricks      | Strong      | Some      | Downward | 2008–2010         |
| Hare        | None        | Strong    | Cyclical | Every 10 years    |
| H02 Cost    | Mild        | Mild      | Upward   | Occasional spikes |
| Gasoline    | Strong      | Some      | Stable   | 2020              |

## Q9: Time Plot Matching

Daily Temperature of a Cow = C

Monthly Accidental Deaths = A

Monthly Air Passengers = D

Annual Mink Trappings = B

## Q11: Google Closing Stock Prices

```{r}
dgoog <- gafa_stock |>
  filter(Symbol == "GOOG", year(Date) >= 2018) |>
  mutate(trading_day = row_number()) |>
  update_tsibble(index = trading_day, regular = TRUE) |>
  mutate(diff = difference(Close))

dgoog
```

-   **Why was it necessary to re-index the tsibble?**

    1.  We had to re-index the *tsibble* because we wanted the **Daily** changes

-   **Plot these differences and their ACF.**
```{r}
autoplot(dgoog, diff)

```



1.  **Do the changes in the stock prices look like white noise?**

    1.  Yes, the changes in stock prices look like white noise.
